---
title: "Themes"
author: "`r paste('Derek Ogle,',format(Sys.time(),'%B %Y'),sep=' ')`"
output:
  xaringan::moon_reader:
    lib_dir: zlibs
    css: ["zlibs/xaringan-themer.css", "hygge", "ninjutsu"]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r purler, eval=FALSE, echo=FALSE}
## Use Knit button to make actual slides.
## Makes (Ctrl-S & Ctrl-Alt-C) a script ... will need to edit.
fnm <- "Lecture_Themes"
FSA::purl2(file.path(here::here(),"modules/Themes",paste0(fnm,".Rmd")),
           moreItems=c("source","setwd","fnm","opts_chunk","flipbookr",
                       "cache","head","include_graphics","options",
                       "splitPerc","leftAssign"))
```


```{r setup, include = FALSE}
source("../zzzflipbook_settings.R")
#!# Loading the tidyverse package (must do every time)
library(tidyverse)
```

class: inverse, center, middle

# Today's Goal

<font size="7">Learn how to provide specific control over all elements of a plot.</font>

---

# Today's Goal

- Demonstrate built-in themes.

--
- Demonstrate how to modify aspects of a plot with `theme()`.

--
- Show how to create your own theme.

--
- Demonstrate how to set a default theme for each plot.

---

class: inverse, center, middle

# What is a Theme?

---

# What is a theme?

- XXX

---

class: inverse, center, middle

# Built-in Themes

---

class: inverse, center, middle

# Controlling Specific Aspects of a Theme

---

class: inverse, center, middle

# Creating Your Own Theme

---

class: inverse, center, middle

# Setting a Theme for all Plots

---


---

class: inverse, center, middle

# Next Time

<font size="7">We will have guest speakers demonstrate case studies.</font>
