---
layout: page
title: Exercise - Annotations
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage (CTRL-S and CTRL-ALT-C)
source(file.path(here::here(),"rhelpers/rhelpers.R"))
fnm <- "CE_Annotations"
dnm <- "modules/Annotations"
modHTML(file.path(here::here(),dnm,fnm))
FSA::purl2(file.path(here::here(),dnm,paste0(fnm,".Rmd")),
           moreItems=c("source","setwd","fnm","dnm","opts_chunk","CTRL-S"))
```
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, fig.width=3, fig.height=3,out.width="60%")
library(tidyverse)
```

----

## Annual Temperature Anomaly
Make the plot shown under ["BONUS - Annual Temperature Anomaly 5" on this page](https://derekogle.com/NCGraphing/modules/Intro2/CE_Temperature.html#bonus-annual-temperature-anomaly-5), but add an appropriate title, subtitle, and caption (and decide if it is worth keeping the x- and y-axis labels).

&nbsp;

## State and No Good Deer Hunting
Make the plot shown under ["State and No Good Deer Hunting 1 on this page](https://derekogle.com/NCGraphing/modules/Univariate_Groups/CE_Wolves.html#state-and-no-good-deer-hunting-1). I have since learned that these data were preliminary and that results from Minnesota hard largely not been entered into the database. Add a note on the **inside** of the plot that explains this. You should include an arrow that appropriately points to the low bar for Minnesota.

&nbsp;

## Total Annual Harvest by County and Year
Legends for very simple plot scan sometimes waste space and be distracting. Such an example may be the legend for the plot shown under ["Total Annual Harvest by County and Year" on this page](https://derekogle.com/NCGraphing/modules/Means/CE_DeerHarvest.html#total-annual-harvest-by-county-and-year). One alternative is to place a small label near the lines in the plot, as shown below. Construct `ggplot2` code to match this graph.

```{r echo=FALSE}
dh <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/DeerHarvest.csv")
cbbPalette <- c("#000000","#E69F00","#56B4E9","#009E73","#F0E442","#0072B2","#D55E00","#CC79A7")

harv5 <- ggplot(data=dh,mapping=aes(x=Year,y=Ttl.All,color=County)) +
  geom_line(size=1) +
  geom_point(shape=21,fill="white",size=2) +
  scale_x_continuous(name="Year",breaks=seq(2010,2018,2)) +
  scale_y_continuous(name="Total Annual Deer Harvest",expand=expansion(mult=0),limits=c(0,8000)) +
  scale_color_manual(values=cbbPalette) +
  theme_bw() + theme(legend.position="none") +
  annotate(geom="label",x=2010,y=2950,label="Ashland",hjust="left") +
  annotate(geom="label",x=2010,y=4600,label="Bayfield",hjust="left",
           color=cbbPalette[2]) +
  annotate(geom="label",x=2010.8,y=7000,label="Douglas",hjust="left",
           color=cbbPalette[3]) +
  annotate(geom="label",x=2010.8,y=900,label="Iron",hjust="left",
           color=cbbPalette[4])
harv5
```

