---
layout: page
title: Exercise - Geoms
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage (CTRL-S and CTRL-ALT-C)
source(file.path(here::here(),"rhelpers/rhelpers.R"))
fnm <- "CE_Geoms"
dnm <- "modules/Intro1"
modHTML(file.path(here::here(),dnm,fnm))
FSA::purl2(file.path(here::here(),dnm,paste0(fnm,".Rmd")),
           moreItems=c("source","setwd","fnm","dnm","opts_chunk","CTRL-S"))
```
```{r echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, fig.show=FALSE, results='hide', message=FALSE, warning=FALSE)
```

----

## Dataset
A very small data set is contained in [Geoms_Example.csv](http://derekogle.com/NCGraphing/modules/Intro1/Geoms_Example.csv). Please load these data[^loaddata] and display the results.[^showdata]

```{r}
#!# Loading the tidyverse package (must do every time)
library(tidyverse)

#!# Load and Prep Data
#!# Set to your own working directory and have just your filename below.
df <- read.csv("http://derekogle.com/NCGraphing/modules/Intro1/Geoms_Example.csv")
df
```

## Geoms I
Create a base plot that maps `x` to the "lon" variable and `y` to the "lat" variable. Add `geom_`s to this base plot to make the eight plots we made during lecture.[^tidyverse]

```{r}
#!# Make the base plot
p <- ggplot(data=df,mapping=aes(x=lon,y=lat))

#!# Adding the "usual" geoms
p + geom_point()
p + geom_line()
p + geom_path()
p + geom_area()
p + geom_polygon()
p + geom_tile()
p + geom_bar(stat="identity")
p + geom_text(aes(label=let))
```

## Geoms II
There are other `geom_`s that we did not discuss during lecture. Of the others, `geom_density_2d()`, `geom_hex()`, `geom_step()`, `geom_rug()`, and `geom_label()` (note that `geom_label()` requires the same `aes()` argument as `geom_text()`. Try adding two of these `geom_`s to your base plot from above. Explain what was plotted (note that you may want to perform an internet search on the `geom_` name for help with what it is plotting).

```{r}
#!# Trying some "new" geoms
p + geom_density_2d()
p + geom_hex()
p + geom_step()
p + geom_rug()
p + geom_label(aes(label=let))
```

## Geoms III
Add one or more of the `geom_` from "Geoms I" above AND one or more of the `geom_`s from "Geoms II" above to create a single plot that is at least somewhat interesting (*with the understanding that it is probably not that interesting yet as we don't have much in our toolbox yet*).

```{r}
#!# Putting geoms together
p + geom_point() + geom_rug()
p + geom_point() + geom_density_2d()
p + geom_point() + geom_hex()
p + geom_line() + geom_text(aes(label=let))
```

[^loaddata]: Download the data and load them from there into R. How to load a CSV file into RStudio is [described in this video](https://vimeo.com/144035138), for which the password is "NCStats" (without the quotes).

[^showdata]: Simply type the name of the object you saved the results in. In the lecture I would just type `dfobj`.

[^tidyverse]: Don't forget to load the tidyverse.

